---
title: "Awe Tracks Prestige: Study 1 Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(lmerTest)
library(sjPlot)
library(piecewiseSEM)
library(lavaan)
library(psych)
```

This document provides the output of the analysis of Study 1 in the paper "Awe Tracks Prestige."

```{r load data}
# C:/Users/danv0000
dp.dat <-  read.csv("C:/Users/danv0000/Dropbox/AwePrestige/DomPres_Study1/dp_inspirers_w.csv")

dp.long <- read.csv("C:/Users/danv0000/Dropbox/AwePrestige/DomPres_Study1/DP_inspirers_long.csv")

# identifying the influence variables
inf <- c("money","guilt","politic","morality","smart",
         "activity","science","news","art","school")
# identifying the emotion variables
emos <- c("amu","ang","awe","comp","cont",
          "disg","emb","enth","fear","guil",
          "love","sad","des","sham","pri",
          "jeal")

# grand mean centering variables for multi-level analysis
dp.long[,c(emos,inf,"dom","pres")] <- scale(dp.long[,c(emos,inf,"dom","pres")], center=T, scale=F)
```

# Demographics

```{r demos}
# Average Age
mean(dp.dat$part_age,na.rm=T)

# SD of Age
sd(dp.dat$part_age,na.rm=T)

# Gender Proportions
prop.table(table(dp.dat$part_sex))

# Ethnicity Proportions
prop.table(table(dp.dat$part_ethn))
```

# Manipulation Check

```{r manip check}
### Manipulation Check
t.test(x=dp.dat$pres_pres, y=dp.dat$dom_pres, paired=T) # for prestige measures
mean(dp.dat$pres_pres) # prestige mean for prestige condition
mean(dp.dat$dom_pres) # prestige mean for dominance condition

t.test(x=dp.dat$pres_dom, y=dp.dat$dom_dom, paired=T) # for dominance measures
mean(dp.dat$pres_dom) # dominance mean for prestige condition
mean(dp.dat$dom_dom) # dominance mean for dominance condition
```

# Relations Between Emotions and Dominance and Prestige

## Awe Alone

```{r awe dp}
# awe predicted by dominance and prestige
awe_dp <- lmer(awe ~ pres + dom + (1|id), data=dp.long)
sjt.lmer(awe_dp, depvar.labels="Awe Response", string.est="Estimate", string.p="p-value",
         separate.ci.col=F, show.std=T, p.numeric=F, pred.labels=c("Prestige","Dominance"))
sem.model.fits(awe_dp)
```

## Fear Alone

```{r fear dp}
# fear predicted by dominance and prestige
fear_dp <- lmer(fear ~ pres + dom + (1|id), data=dp.long)
sjt.lmer(fear_dp, depvar.labels="Fear Response", string.est="Estimate", string.p="p-value",
         separate.ci.col=F, show.std=T, p.numeric=F, pred.labels=c("Prestige","Dominance"))
sem.model.fits(fear_dp)
```

## All Emotions Predicting Prestige and Dominance

```{r all emos}
pres_emos <- lmer(pres ~ amu + ang + awe + comp + cont + disg + emb +
                    enth + fear + guil + love + sad + des + sham + 
                    pri + jeal + (1|id), data=dp.long)
dom_emos <- lmer(dom ~ amu + ang + awe + comp + cont + disg + emb +
                   enth + fear + guil + love + sad + des + sham + 
                   pri + jeal + (1|id), data=dp.long)

sjt.lmer(pres_emos, dom_emos, depvar.labels=c("Prestige","Dominance"), string.est="Estimate", string.p="p-value", separate.ci.col=F, show.std=T, p.numeric=F, 
         pred.labels=c("Amusement","Anger","Awe","Compassion","Contentment","Disgust",
                       "Embarrassment","Enthusiasm","Fear","Guilt","Love","Sad",
                       "Desire","Shame","Pride","Jealousy"))

sem.model.fits(pres_emos)
sem.model.fits(dom_emos)
```

# Analysis of Influence

## Parallel Analysis for Influence Variables

```{r influence fa}
# creating a data set with just influence variables
inf.dat <- dp.long[,inf]
# conducting the parallel analysis
fa.parallel(inf.dat)
# since the items appear to be uni-dimnesional, I create a single influence scale
dp.long$inf <- rowMeans(dp.long[,inf], na.rm=T)
```

## Predicting Influence from Dominance and Prestige

```{r inf dp}
# effect of prestige alone
inf_p <- lmer(inf ~ pres + (1|id), data=dp.long)
summary(inf_p)
sem.model.fits(inf_p)

# effect of dominance alone
inf_d <- lmer(inf ~ dom + (1|id), data=dp.long)
summary(inf_d)
sem.model.fits(inf_d)

# effect of both dominance and prestige
inf_dp <- lmer(inf ~ dom + pres + (1|id), data=dp.long)
summary(inf_dp)
sem.model.fits(inf_dp)
```

## Predicting Influence from Awe

```{r inf awe}
inf_awe <- lmer(inf ~ awe + (1|id), data=dp.long)
summary(inf_awe)
sem.model.fits(inf_awe)

# effect of awe, controlling for prestige
inf_p_awe <- lmer(inf ~ pres + awe + (1|id), data=dp.long)
summary(inf_p_awe)
sem.model.fits(inf_p_awe)
```

## Mediation Analysis: Awe Only

```{r med awe only}
dp.dat$dom_inf <- rowMeans(dp.dat[,c("dom_money", "dom_guilt", "dom_politic", "dom_morality", "dom_smart", 'dom_activity', "dom_science", "dom_news", "dom_art", "dom_school")], na.rm=T)

dp.dat$pres_inf <- rowMeans(dp.dat[,c("pres_money", "pres_guilt", "pres_politic", "pres_morality", "pres_smart", 'pres_activity', "pres_science", "pres_news", "pres_art", "pres_school")], na.rm=T)

awe_med_p <- '
pres_awe ~ a*pres_pres

pres_inf ~ b*pres_awe + cp*pres_pres

ab := a*b
tot := a*b + cp
'

awe_med_p.fit <- sem(awe_med_p, dp.dat, se="bootstrap")
summary(awe_med_p.fit, standardized=T, rsquare=T)

awe_med_d <- '
dom_awe ~ a*dom_pres

dom_inf ~ b*dom_awe + cp*dom_pres

ab := a*b
tot := a*b + cp
'

awe_med_d.fit <- sem(awe_med_d, dp.dat, se="bootstrap")
summary(awe_med_d.fit, standardized=T, rsquare=T)
```
